Dynamics of international giving: Exploring individual donor preferences

### 2020-06-16 manuscript/manuscript.Rmd
Suparna initial R&R edits

---
title: "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  email: suparna.chaudhry@cnu.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on NGOs have become increasingly pervasive across the globe. With reduced official aid and foundation funding available for NGOs, and an absence of philanthropic culture in many countries across the Global South, various NGOs are seeking funds from individual private donors in Western countries. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? While existing research has shown that donors rely on various heuristics from social networks and organization-level attributes, we argue that domestic political environments of NGO host countries can also serve as an important structural-level heuristic that can influence donor behavior. Using a survey experiment, we find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We discuss the implications of our results on the sustainability of organizations working in repressive countries.
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of Political Science, Christopher Newport University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of Political Science, Christopher Newport University, Newport News, VA 23606, 757-594-7821. Contact: suparna.chaudhry@cnu.edu
word-count: "4,999"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Following the 2011 Egyptian uprising, just as domestic non-governmental organizations (NGOs) began to take advantage of new levels of civic freedoms, successive Egyptian governments passed laws restricting these groups from receiving foreign funding or engaging in political advocacy. Egypt is not alone in this endeavor—India, Bolivia, Hungary, Russia, and others have also imposed funding restrictions on NGOs [@Carothers:2015; @DPR:2015]. Consequently, official aid flows and grants from large foundations to NGOs have considerably decreased in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018].

In response to this shift in the NGO funding landscape, organizations have increasingly turned to *foreign* individual donors to offset the loss of governmental and foundation funding, [@BanksHulmeEdwards:2015] as funding from *domestic* donors in repressive countries is unreliable, inconsistent, and frequently prohibited. Citizens in the Global South often cannot support local NGOs, may prefer channeling funds to groups working on innocuous issues such as health and education over more contentious issues, or may be deterred by dismal tax incentives or a general fear of retribution [@DPR:2015; @RPC:2016; @Brechenmacher:2017]. NGOs have thus increasingly sought out funds from *foreign* individual donors in countries with institutions more amenable to philanthropy. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017].

In the face of increased crackdown on NGOs globally, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate? Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974]. While much of this research has focused on heuristics from social networks and organizational practices [@CrosonShang:2011; @Tremblay-BoirePrakash:2017], in this research note we argue that domestic political environments of NGO host countries can serve a structural heuristic that influences individual donor decisions. When donors are provided salient information about crackdown on NGOs and its impact on their activities, that information can serve as a heuristic of organizational deservingness, and consequently influence donor preferences and behavior. We use a survey experiment to measure the effect of crackdowns on stated donor preferences and find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate substantially more to legally restricted NGOs. This effect persists when mediated by other organization-level heuristics like NGO issue area and its main funding sources.

This research note makes two important contributions. While recent work has looked at how crackdowns on NGOs affect official aid [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how this affects private aid. Further, while there is a robust existing literature on the individual-level determinants of philanthropy, far less is known about its structural-level determinants, particularly towards international causes, and how these factors interact with organizational attributes to signal NGO deservingness.

Below, we summarize existing research on the determinants of giving to NGOs and lay out our expectations regarding how crackdowns on NGOs may affect private philanthropy. We then describe our survey and present the results. We conclude with implications for NGOs working in repressive countries and lay out questions ripe for future research.

# Theorizing the effects of crackdown on philanthropy

Substantial research on charitable giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011], and more recently, the effect of social networks on philanthropy [@HerzogYang:2018]. However, research on individual donor behavior has overwhelming examined giving to local—rather than international—causes. While most research on giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014], we explore private philanthropy because individuals rely on signals and heuristics that may differ from the benchmarks used by large governmental and private donors.

The decision to donate to an NGO is rarely rational. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. Similarly, the effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009].

@@

> **H~1a~**: If NGOs face legal crackdowns abroad, then individual donors will be more likely to donate to them.
>
> **H~1b~**: If NGOs face legal crackdowns abroad, then individual donors will donate more to them.

@@

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences—can serve as an organization-level heuristic.[^targeting] NGOs that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. NGOs working on these less contentious issues can also frame their activities as in need of greater support—the perception of need, whether through personal experiences or media framing, strongly influences individual decisions to donate to relief efforts [@BekkersWiepking:2010, 6–8].

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. Thus, donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes. As such, we hypothesize that donors will be more likely to donate to less contentious NGOs because these are seen as uncontroversial, apolitical, directed towards the most deserving of need, and are more capable of producing short-term quantifiable results. We use humanitarian NGOs as a proxy for NGOs working on non-contentious issues and human rights NGOs as a proxy for contentious issues.

> **H~2a~**: If legally besieged NGOs work on humanitarian issues, then individual donors will be more likely to donate to them.
>
> **H~2b~**: If legally besieged NGOs work on humanitarian issues, then individual donors will donate more to them.

@@

> **H~3a~**: If legally besieged NGOs do not receive substantial funding from government sources, then individual donors will be more likely to donate to them.
>
> **H~3b~**: If legally besieged NGOs do not receive substantial funding from government sources, then individual donors will donate more to them.

@@

# Discussion: Do crackdowns on NGOs affect individual philanthropy?

@@

This research note should be viewed as an exploratory study, and it raises numerous questions for future research. These include determining the extent to which these preferences translate into behavior—here, we have only asked respondents to make a hypothetical donation, but preferences may not reflect actual behavior. Future research can also test the generalizability of our findings to other issue areas and political contexts.

These results have several implications for policy. NGOs may benefit from publicizing when they are targets of government crackdown. While this may not be enough to sway the average U.S. citizen, our results show that communicating this information can convince already-likely donors to donate more to besieged groups. Further, if NGOs signal to donors that they receive a majority of their funding from non-governmental sources, they may be able to convince average donors that their contributions might help the NGO resist the crackdown. Moreover, since private funders typically tend to be less interested in the domestic politics of recipient countries compared to governmental funders, relying on private donations is less likely to attract the ire of authoritarian governments fearful of foreign intervention.

@@

>>>>

---
title: "Dynamics of international giving: Exploring individual donor preferences"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Lewis \& Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on NGOs have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private funding flows. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO regulatory environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and its main funding sources. We conclude by discussing the implications of our results on the scholarship and practice of nonprofits using different framing appeals to increase fundraising at a time when traditional funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "FILL"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Over the last three decades, governments across the world have sought to limit the work of nonprofits - a phenomenon knowns as `closing civic space' in the policy arena [@Carothers:2015; @DPR:2015]. Government have enacted laws creating barriers to advocacy, funding, and entry, adversely affecting the work of thousands of NGOs, which often receive a large amount of resources from donors across the world. This is particularly the cause in countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency and donors seek out NGOs for projects (Dietrich 2013). Consequently, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on public funding flows, we know relatively little about how this crackdown affects private donors, especially individual donors' willingness to donate.

Charitable-giving to nonprofits working in international affairs has been steadily increasing over the past two decades, with a noticeable increase over the last decade. Between 2014 and 2016, giving to international affairs increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion - or approximately 70\% of total giving to international affairs. This further increased to $427.71 billion in 2018 - 85.7% of which was through individuals [givingusa:2019]. Given that a majority of this research is restricted to elite, high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of individual donors. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about regulatory environments abroad, as well as organizational-level attributes of nonprofits, such as their main focus areas and sources of funding affect donors' willingness to donate internationally?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974; @CrosonShang:2011; @Tremblay-BoirePrakash:2017], We leverage an online survey experiment using Amazon's Mechanical Turk to assesss how different structural and organizational-level heuristics affect individual donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: NGO issue area and its main funding sources. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown - with donors showing an increased willingness to not just donate to them, but also to donate more to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to governments, and that such organizations need greater support when facing stricter regulatory environments.

Our study makes a significant contribution to the scholarship and practice on dynamics of individual-giving and indicates the promise and limits of different framing appeals by nonprofits on individuals' willingness to donate. First, while factors giving shaping individual-giving to domestic causes have been thoroughly explored in existing literature, the dynamics of internatoinal giving are under-examined in the U.S. literature on individual donor preferences.[^exception] It is even more essential to understand as philanthropy to international causes can be harder to motivate because of the number of recipients are larger and far removed from the donor [@CasaleBaumann:2015, 100].

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to infleunce perceptions of behavior, is particular important in this regard.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to donors considering international charitable-giving. Our study looks at the impact of political frames such as regulatory crackdown, and how it interacts with organizational attributes to signal NGO deservingness. In doing so, it can help nonprofits frame their appeals.

CHANGE
Below, we summarize existing research on the determinants of giving to NGOs and lay out our expectations regarding how crackdowns on NGOs may affect private philanthropy. We then describe our survey and present the results. We conclude with implications for NGOs working in repressive countries and lay out questions ripe for future research.

#Philanthropy towards international nonprofits

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though our focus in this paper is one private funding, it is essential to differentiate between the two sources. Public aid (more commonly known as traditional donor aid) channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds, which are comparatively newer international donors compared to traditional donor aid, can come from foundations, private corporations and individuals.

In this paper, we specifically look at individual-level determinants to foreign giving. As described in the introduction, this is because [ADD PART ABOUT ANY PUBLIC FIGURES AND HOW IT DECLINING] private-giving to nonprofits working in international affairs has been consistently increasing over this past decade.[^internationalaffairs] Moreover, private giving by individuals has far outstripped giving by private foundations - in 2018, charitable-giving by individuals amounted to $292.09 billion, compared by $75.86 billion by foundations and $20.05 billion by corporations. Thus, giving by foundation represented 18\% of all charitable-giving in 2018, which was not only lesser than the proportion of giving by private individuals, but also giving by Bill and Melinda Gates Foundation accounted for more than half of it [givingusa:2019].

Given these substantive amounts, understanding individual-level determinants of giving to international nonprofits is important to investigate as individual private donors may not necessarily mirror public donor choices. While there is evidence that citizens generally support the goals of aid agencies (Milner and Tingley 2010), surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, foreign aid and democracy aid is often considered unresponsive to developments on the ground [@brainard2007, @Carothers:2015] and public aid decisions - especially on final recipients are not easily accessible. There may also be heterogeneity within how donor governments allocate funds. Emerging donor government differ from established donors in that they demand fewer human rights, governance or environmental conditions preferred by Western donors [@fenglerkharas].

More recent research on the allocation of private aid flows has also found little support for the idea that individual donor behavior mimics that of official donors. Desai and Kharas (2018) find that unlike official donors  whotend to channel funds to countries with better institutional quality (control of corruption) and as a performance reward (Bermeo 2011, Dietrich ad Wright 2017), individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to groups in countries experiencing a humanitarian crisis - in particular, those driven by natural disasters. They go on to conclude that compared to donor agencies, "private donors respond to different project and country characteristics." (517) Given these above reasons, decision-making by the average individual donor may therefore not have the same logic as donor governments and should be investigated on its own terms.

Understanding motivations of individual private donors is also essential because private funding flows towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. Prior research has demonstrated that restrictions on cross-border funding has reduced official aid and official development assistance to repressive countries (CITE). But data from the Foundation Center shows no correlation between amount of funding flows from US private foundations to a given country and that country's score on the Index of Philanthropic Freedom. While this index is comparatively new and data is available only 2015 onwards (CHECK), preliminary evidence shows that, "US foundations remain committed to channeling funds to countries where the legal environment is growing increasingly unfriendly." As such, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to countries with restrictive legal environments.

##Key drivers of individual-level philanthropy

Substantial research on charitable giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011]. However, research on individual donor behavior has overwhelming examined giving to local—rather than international—causes. This literature has primarily focused on the importance of X main factors, especially when considering charitable behavior towards international causes: (1) the role of social capital and social networks; (2) the role of individual experiences such as higher education and levels of religiosity; and (3) third party certifications.

One major area of inquiry in philanthropic studies looks at the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit." (Putnam 1997, 67). While quite broad, the literature further concentrates on a specific kind of social capital-associational capital-which refers to, "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations. This is because volunteering exposes individuals to the need for donations to charities and nonprofits and made them feel more compelled to do so [@hossainlamb2012].

The role of social networks is also important in this regard. Looking at relational configurations in social networks, Herzog and Yang find that having both a giver and a solicitor in one's social network increase the likelihood of that individual's charitable-giving.

Individual experiences such as access to higher education and increased religiosity can also impact philanthropic giving. Higher education fosters prosocial motivations and also brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. This in turn has been shown to be particularly vital in giving to international causes. Additional education has also been shown to increase abstract thinking [@wiepkingmass, 1978]. This is particular important for donating to international relief organizations than donating to domestic organizations [@micklewright2009].

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable-giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015].

Finally, information on organizational characteristics and third party certifications can also impact levels of individual charitable-giving, through the results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits (Szper and Prakash 2011).

While a robust literature as shown above looks at the drivers of individual-level philanthropy, it overwhelmingly focusses on giving to local rather than international causes. This focus on local or domestic causes means that that the precise causal mechanisms through which cross-border charity can be motivated and mobilized is still under-examined (Desai and Kharas, 517). This problem is further compounded by the fact that there is greater paucity of data and diversity of sources regarding international-giving, which further complicates a systematic examination.

Indeed, most research on individual-giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014].[^highnetworth] And recent research has on the average donor finds that, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad." (TB Prakash 2017, 644) And as explained above in section X, private donors do not necessarily mirror the choices made by public donors and that individuals rely on signals and heuristics that may differ from the benchmarks used by large governmental and private donors.

These signals and heuristics, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may be therefore more reliant on framing. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

#Theorizing the effects of political and organizational frames on individual-giving

@@

> **H~1a~**: If donors learn that NGOs face legal crackdowns abroad, then individual donors will be more likely to donate to them.
>
> **H~1b~**: If donors learn that NGOs face legal crackdowns abroad, then individual donors will donate more to them.

@@

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences, can also serve as an organization-level heuristic.[^targeting] NGOs that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. Previous research has shown that that public donors do indeed differentiate between NGO issue areas, especially when these organizations are working in repressive contexts. In particular, the Organisation for Economic Cooperation and Development (OECD) donors respond to repressive NGO environments, especially those containing barriers to advocacy, by decreasing funds for groups working on politically sensitive causes such as anti-corruption initiatives, elections, human rights, legal reform and security sector reform, and instead increasing funding for groups working on relatively tamer causes such as health, education and agriculture [CITE US].

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. There is evidence that this a concern for private donors. In a survey of women’s NGOs across the globe conducted by Mama Cash and Urgent Action Fund, many NGOs argued that private donors (especially foundations) were withdrawing from funding sensitive issues in repressive contexts, which in turn was hurting women and trans rights organizations. Some private funders, these groups argued, were also more likely to fund well-established groups that were more likely to survive crackdown.

However, compared to our existing knowledge on public donors and foundation donors, we lack knowledge about how individual donors respond to frames emphasizing one nonprofit issue area over another, especially a contentious area over a tamer one. NGOs working on tamer issues such as health, education, disaster relief, and humanitarian aid can often frame their activities as in need of greater support—the perception of need, whether through personal experiences or media framing. This may strongly influence individual decisions to donate to relief efforts [@BekkersWiepking:2010, 6–8]. Individual donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes.

As such, we hypothesize that donors will be more likely to donate to less contentious NGOs because these are seen as uncontroversial, apolitical, directed towards the most deserving of need, and are more capable of producing short-term quantifiable results. We use humanitarian NGOs as a proxy for NGOs working on non-contentious issues and human rights NGOs as a proxy for contentious issues.

> **H~2a~**: If donors learn that legally besieged NGOs work on humanitarian issues, then individual donors will be more likely to donate to them.
>
> **H~2b~**: If donors learn that legally besieged NGOs work on humanitarian issues, then individual donors will donate more to them.

@@

> **H~3a~**: If donors learn that legally besieged NGOs do not receive substantial funding from government sources, then individual donors will be more likely to donate to them.
>
> **H~3b~**: If donors learn that legally besieged NGOs do not receive substantial funding from government sources, then individual donors will donate more to them.

@@

 I would encourage you to revisit the hypotheses from the findings. Which hypotheses were supported, and which were not? Why does it matter whether NGOs are advocating human rights or provide humanitarian assistance, and how much government funding they receive? How did participants interpret the term government funding?

## Discussion
While the results from the survey experiment above allows to hold the organization constant and test for specific hypotheses, there are a few caveats when interpreting these results, generalizing them to other contexts, and considering applied lessons for fundraisers. First, the experiment measures preferences (willingness to donate) rather than behavior (actual donations). While the former may not always neatly translate into the latter, there is evidence that individuals’ willingness to donate as measured in the experiment translates well into behavior. This is because the wording of the survey experiment uses an informational frame, where the focus is on educating the rather by presenting them with facts. Prior research has shown that informational frames (as opposed to the use of personal frames, which draw the audience’s attention to the plight of an individual, or motivational frames, which are meant to motivate individuals to act by creating feels of agency and efficacy) are important in generating donations (McEntire, Lieby, Kraine 2015).[^frametypes]

Moreover, research has also found that in the case of campaigns for human rights nonprofits, informational frames are a necessary precondition for getting people to act, and people are more likely to donate and undertake other actions when these informational frames are combined with personal frames (McEntire, Lieby, Kraine 2015).[^personalframes] This suggests that if governments attempt to repress international human rights nonprofits, then these organizations can potentially increase their resource flows by using a combination of these two frames. The survey instrument used above shows substantive increase in donors' willingness to donate to human rights groups facing crackdown, especially those that are privately-funded, and this result emerges from the use of just informational frames. If combined with personal frames, research indicates these donations may further increase.

It is also worth noting that there is evidence, at least domestically, that restrictions on nonprofit activities spur changes in private donors’ actual giving patterns (and not merely their preferences). For instance, following the 2016 election, when Texas pledged to remove Planned Parenthood from its Medicaid program and a number of states proposed new restrictions on their work, donations to the organizations increased by over 40 times the usual amount.

your convenience sample captures those relatively more likely to donate to charities, which perhaps allows you to have more specific lessons for fundraisers.

Finally, given that the survey was administered to a convenience sample of Americans, it raises questions about external validity. 1. ADD HOW RESULTS MAY DIFFER FOR REPRESENTATIVE SAMPLE OF AMERICANS. 2.  Would a sample of Europeans for instance be so skeptical towards government funding?

# Conclusion

@@

This article should be viewed as an exploratory study. While the survey experiment allows us to hold the organization constant and test for specific hypotheses, there are limits to its external validity. Future research should look to testing the generalizability of our findings to individual donors in other countries, as well as determining the extent to which these preferences translate into behavior. It is also important to explore among thsoe donors who give to international causes - explore the extent to which individuals are likely to give to domestic nonprofits abroad, compared to international nonprofits working in those countries.

These results also have implications for nonprofit fundraising. NGOs may benefit from publicizing when they are targets of government crackdown and including that information in their framing appeals. While this may not be enough to sway the average U.S. citizen, our results show that communicating this information can convince already-likely donors to donate more to besieged groups. Further, if NGOs signal to donors that they receive a majority of their funding from non-governmental sources, they may be able to convince average donors that their contributions might help the NGO resist the crackdown.

@@

[^internationalaffairs]: Giving USA defines the sector of international affairs to include any nonprofit organization working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs. This classification is based upon the National Taxonomy of Exempt Entities Core which codes groups based upon their field of interest.

[^exceptions]: For exceptions, see @micklewright2009 on UK, @Rajan:2009 on Canada, @wiepking2010 on Netherlands, @CasaleBaumann:2015 on the USA.

[^highnetworth]: High-net worth households are defined as having greater than $200,000 in annual household income and/or with more than $1,000,000 in assets, excluding primary residence.

[^framing]: Framing is the process through which actors present information to influence perceptions or behavior. Previous research has shown that which frames are employed can have a significant effect on individuals' attitudes and behaviors [Benford and Snow 2000, Chong and Druckman 2007].

[^frametype]: For more on motivational and personal frames, see Benford and Snow 2000, Gamson 1995

[^personalframes]: Personal frames seek to humanize victims and entice action by invoking an emotional reaction. Research has found that people are more likely to be generous when donating to a human rights campaign with a single victim, as opposed to a statistical group of victims (Slovic 2007).

[^NGO]: Nonprofit organizations (which we inter-changeably use with NGOs) are organizations for a public or mutual benefit other than generating profit for the owners or investors (Salamon 1998). NGOs can be either domestic i.e. an NGO that operates entirely in a single state, or international, more commonly known as INGOs. INGOs or international nonprofits, "operate independently of governments, are composed of members from two or more countries, and are organized to advance their members' international goals and provide services to citizens of other states through routine transactions with states, private actors and international institutions." (Tarrow 2001) Unlike domestic NGOs, INGOs often work in countries outside of their main country of origin or headquarters.
